Depression and suicidality as evolved credible signals of need in social conflicts

### 2021-06-11 Paper.rmd
Minor edits, new aversiveness subsection

In the bargaining framework, suicidality is conceptualized as putting all future contributions to cooperative endeavors with social partners at risk with some low but non-zero probability, credibly signaling low valuation of current circumstances. On this view, most suicides deaths, especially in young, physically healthy individuals, would therefore be the inevitable consequence that some individuals lose this gamble. These and similar ideas [@andrews2006; @HagenEtAl2008; @nock2008; @Stengel1956; @Rosenthal1993] suggest that depression symptoms that put one's value to others at risk, such as loss of interest and suicidality, are credible signals of need.

@@

Anger is aversive to targeted social partners, yet it is probably an adaptation that exploits advantages in physical or social formidability to force concessions from them [@SellEtAl2009]. Aversive responses to depressive and suicidal bargaining are likewise expected throughout the process, encouraging concessions by interdependent social partners, who in turn signal the costliness of increasing their support [@Hagen2003; @SymeHagen2021].

>>>>

In the bargaining framework, suicidality is conceptualized as putting all future contributions to cooperative endeavors with social partners at risk with some low but non-zero probability, credibly signaling low valuation of current circumstances. On this view, most suicides deaths, especially in young, physically healthy individuals, would therefore be the inevitable consequence that some individuals lose this gamble.

@@

## Aversiveness is a feature of depression, not a bug

Under the bargaining model, aversive responses to depressive and suicidal bargaining are expected throughout the process, encouraging beneficial concessions by interdependent social partners with whom one is in conflict, who in turn signal the costliness of increasing their support [@Hagen2003; @SymeHagen2021]. This predicted pattern is quite similar to anger, which is aversive to targeted social partners, yet is probably an adaptation that exploits advantages in physical or social formidability to force beneficial concessions from them [@SellEtAl2009].

In summary, we propose that depression symptoms that put one's value to others at risk, such as loss of interest and suicidality, are aversive but adaptive signals of need [for similar views, see @andrews2006; @HagenEtAl2008; @nock2008; @Stengel1956; @Rosenthal1993; @firth1936; @Firth:1961; @farberow1961].
